---
title: "Working Document"
author: "Chase Jensen"
date: "2023-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

# 1. Business Problem Statement


With 10% of their business coming from the performance of local businesses, 
Swire Coca-cola is looking for a way to determine which potential customers will
be the most successful. With a method in place to predict the success of 
customers, Swire will be able to select the best businesses to sell too. 
This will allow for Swire to set prices that benefit both themselves and the 
businesses and also maximize potential ROI. There are not any immediate success 
metrics that we can measure, so determining success will take some time. 

However, after a certain amount of time, success can be determined by the 
revenue and traffic the store brings in. 
The approach will revolve around analyzing the customer sentiment 
(online reviews), sales volume, and years of operation of other Swire businesses 
in a selected area. The analysis will ultimately result in a predictive model 
that will identify favorable business targets. Additionally, providing an 
initial list of potential targets for Swire. The project will be completed by 
Hobeen Moon, Joey Chen, Nik Sassen, and Chase Jensen with a delivery date of 
April 12th. Specific milestones will include the identification of the business 
problem, exploratory data analysis, building of the model, evaluation of the 
model, and final results in the form of a presentation.

# 2. Analytic Objective


The objective of this analysis is to is to create a model that is able to 
accurately predict the probability that a potential customer will be successful.
While the actual method of model creation is still to be determined, the 
outcome will be the same. We will determine this a successful analysis if we 
can create a model that can determine a likelihood for a business to succeed 
with high accuracy that can then be given over to the client for implementation.

# 3. EDA Questions


What do the distributions of sales numbers look like; namely Revenue, Discount, and Quantities?
What customer demographics are the most common? Where are they most likely to be located? What types of businesses are there? 
What are these customers purchasing? How often are they purchasing?


# 4. Data Exploration

```{r Data Reading and Libraries, message=FALSE, warning=FALSE}
library(readxlsb)
library(readxl)
library(tidyverse)
library(ggplot2)
library(ragtop)
library(ggmap)

dataDict <- read_excel("Data Dictionary.xlsx")
swireTerm <- read_excel("Swire Terminology.xlsx")
custData <- read_excel("FSOP_Customer_Data_v2.0.xlsx")
salesData <- read_excel("FSOP_Sales_Data_v2.0.xlsx")

```


Customer Data Feature Engineering
```{r FE Cust Data}
# Creating a column for State
custData$STATE <- substr(custData$SALES_OFFICE_DESCRIPTION, 
                         nchar(custData$SALES_OFFICE_DESCRIPTION) - 1, 
                         nchar(custData$SALES_OFFICE_DESCRIPTION))
```




# EDA - Store Data

Looking at the structure of the data
```{r EDA for Customer Data}
head(custData)
str(custData)

# Count of NA's
sapply(custData, function(x) sum(is.na(x))) 
# There are only 2 NA's in the entire Customer Dataset, namely County info


# Let's make some tables
table(custData$CUSTOMER_ACTIVITY_CLUSTER_DESCRIPTION) # Customer Activities
table(custData$CUSTOMER_TRADE_CHANNEL_DESCRIPTION) # Trade
table(custData$CUSTOMER_TRADE_CHANNEL_DESCRIPTION2) # I believe that this is the same as above
table(custData$CUSTOMER_SUB_TRADE_CHANNEL_DESCRIPTION) # Further breakdown of trade
table(custData$MARKET_DESCRIPTION) # Mostly On-Premise
table(custData$COLD_DRINK_CHANNEL_DESCRIPTION) # Mostly eating and drinking, some destination
sort(table(custData$STATE), decreasing = TRUE)

```
Count of NA values:
```{r Customer NAs}
sapply(custData, function(x) sum(is.na(x)))
# Only 2 NA's in County 

```


Let's make some plots

```{r Customer Plots}
ggplot(custData, aes(CUSTOMER_ACTIVITY_CLUSTER_DESCRIPTION)) + geom_bar() + 
  coord_flip() + labs(title = "Count of Customer Activity Cluster")
# Majority of businesses are "Eating and Drinking"

ggplot(custData, aes(CUSTOMER_TRADE_CHANNEL_DESCRIPTION)) + geom_bar() + 
  coord_flip() + labs(title = "Count of Customer Trade Channel")
# The two largest trade channels are Quick Service and Full Service Restaurants

ggplot(custData, aes(CUSTOMER_SUB_TRADE_CHANNEL_DESCRIPTION)) + geom_bar() + 
  coord_flip() + labs(title = "Count of Customer Sub-Trade Channel")
# There are a TON of sub-trades. May be valuable to a model, but could 
# need consolidation of sub-trades.


ggplot(custData, aes(MARKET_DESCRIPTION)) + geom_bar() + coord_flip() + 
  labs(title = "Count of Market Description")
# Vast Majority is FSOP

ggplot(custData, aes(COLD_DRINK_CHANNEL_DESCRIPTION)) + geom_bar() + 
  coord_flip() + labs(title = "Count of Cold Drink Channel")
# Majority is eating and drinking, but a good amount of other channels as well

ggplot(custData, aes(STATE)) + geom_bar() + coord_flip() + 
  labs(title = "Count of State")
# This is an attempt to visualize where the businesses were, could be fixed

# Trying a lat-long graph
library(mapview)

ggplot(custData, aes(x = GEO_LONGITUDE, y = GEO_LATITUDE, color = STATE)) + 
  geom_jitter() + xlim(-130, -70) + ylim(30, 50)

mapview(custData, xcol = "GEO_LONGITUDE", ycol = 
          "GEO_LATITUDE", crs = 4269, grid = FALSE)
```

Looking at Sales Data

```{r EDA for Sales Data}
table(salesData$BEV_CAT_DESC) # Overwhelmingly CORE SPARKLING
table(salesData$CALORIE_CAT_DESC) # Regular calorie has the edge over low calorie
table(salesData$PACK_TYPE_DESC)
table(salesData$PACK_SIZE_SALES_UNIT_DESCRIPTION) # That's a lot of packages
# This really could be simplified.

# Numeric Data
summary(salesData$PHYSICAL_VOLUME) # Right Skewness
summary(salesData$INVOICE_PRICE) # Right Skewness
summary(salesData$DISCOUNT) # Right Skewness
summary(salesData$NUM_OF_TRANSACTIONS) # Right Skewness
# Potentially log-transform

# NA's
sapply(salesData, function(x) sum(is.na(x)))
# A lot more NA's here.
# 73,623 for both Beverage and Calorie Category



```




Plots for Sales Data

```{r Sales Data Plots}
ggplot(salesData, aes(BEV_CAT_DESC)) + geom_bar() + coord_flip() + 
  labs(title = "Count of Beverage Category")
# Majority is Core-Sparkling, but a good spread outlside of that

ggplot(salesData, aes(CALORIE_CAT_DESC)) + geom_bar() + coord_flip() + 
  labs(title = "Count of Calorie Category")
# Majority regular calories, but low calorie has a sizeable chunk

ggplot(salesData, aes(PACK_TYPE_DESC)) + geom_bar() + coord_flip() + 
  labs(title = "Count of Package Type")
# Mostly plastic bottles, aluminium cans, and bag in box
# Potentially an area to consolidate. Lots of plastics.

ggplot(salesData, aes(NUM_OF_TRANSACTIONS)) + geom_histogram()

ggplot(salesData, aes(NUM_OF_TRANSACTIONS)) + 
  geom_histogram() + 
  xlim(0, 250) + 
  ylim(0, 200000) +
  labs(title = "Distribution of Number of Transactions")
# A bit of a better look, cutting out the 727 accounts with more than 250 transactions
  

```


```{r}

length(unique(salesData$CUSTOMER_NUMBER_BLINDED)) # 40390 unique customers
mean(salesData$NUM_OF_TRANSACTIONS)





```



Initial Notes:

Probably will want to rename variables
Some variables will be able to be dropped


There are 53,373 instances of negative physical volume
Similarly, there are 72,657 instances of zero physical volume


There are 132,977 instances of zero invoice price and 4,744 instances of 
negative invoice price.

Possibly Omit these as damaged goods/returns


There are 4,199 instances of a negative discount? I'm unsure what that means.



# 5. Results and Findings

The results of this EDA give us some interesting findings, but also raises some
other questions that will need to be addressed. First and foremost, the graphs
and tables give us an understanding of what the majority classes are and what
trends follow those classes. We are finding that the majority of customers
are either quick service or full service restaurants purchasing plastic bottles,
cans, and bag-in-box products (which serve soda fountains).

Questions arise when looking at the sales data. For example, we have several
metrics that, from a logical perspective, should not be negative but are indeed 
negative, including discount and invoice price. Does this mean there is price
inflation and/or sales back to the client? This would need some more in-depth
analysis and potential discussions with the client.


The next steps of this analysis would be to being looking for correlations and
other relationships within the data. This would clear a way for us to begin to
create a model to predict customer success.


# 6. Ethical Considerations

In terms of ethical considerations, there are a few areas that are important to 
keep an eye on. While we do not have customer names or addresses, we do have 
enough information to determine the businesses we are analyzing. Keeping this
information confidential and safe is crucial to the security of the client. 